Parameters associated with Injuries in the NFL

### 2024-07-13 products/manuscript/Manuscript.qmd
Project submission #4

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
library(tidyverse)
```

@@

For years, the NFL has been making changes in the game in an attempt to prevent player injuries. These changes range from rule modifications to investment in advancements to the players pads and helmets. Hopefully, the findings from this project can help inform players, coaches, and who knows, the NFL about the metrics for a higher risk in injury.

@@

Three data sets have been identified for this project. The primary data set ("Injury Player Data" obtained from a Github repository contains 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season. Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics. Two other data sets were obtained from Kaggle. The first is a "cumulative play data" containing information about every play ran in the 2019-2020 NFL season (42,186 observations). Information such as the teams playing, play description, and time in the game when it was ran. The last data set is the "player demographic data" containing information about all NFL players in the 2016-2019 seasons (11,145 observations). This data set has player metrics such as height, weight, date of birth, etc.

@@

*****To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].******

@@

The primary idea behind the analysis for this project is to look at various parameters when an injury occurred and compare them to the rest of the NFL. If the parameters are related to the play itself, the paramter in the injruy data set will be compared to cumulative play data set. Likewise, if the paramter is related to the injured player themselves, this paramter will be copmared to the cumulative player data set. If there is a statistical difference between paramters, we can likely draw some conclusion about a parameter causing injuries.

@@

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

@@

As for the "cumulative player data" set, empty columns will be removed and 0's in the down column will be converted to "NA" values which indicate kickoffs & PATs. Lastly, a lot of unnecessary plays are included in the "cumulative play data" set such as end of quarter/games, 2-minute warnings, and timeouts so these observations can be filtered out.

@@

```{r}
hist(injuries$qtr,main="Injuries by Quarter", ylim=c(0,550), xlim=c(1,5),xlab="Quarter",labels =T)

```
```{r}
hist(pbp$Quarter, breaks=6,main="Total Plays by Quarter", ylim=c(0,11200), xlab="Quarter",labels =T)

@@

```{r}
hist(injuries$down, breaks=5,main="Injuries by Down", ylim=c(0,700),xlim=c(1,4),xlab="Down",labels =T)
```

```{r}
hist(pbp$Down, breaks=6,main="Total Plays by Down", ylim=c(0,14900),xlim=c(1,4), xlab="Down",labels =T)
```

@@

```{r}
hist(injuries$height, breaks=10,main="Injuries by Height", ylim=c(0,250), xlab="Height",labels =T)

```
```{r}
hist(players$height, breaks=10,main="Total Players Height", ylim=c(0,250), xlab="Height",labels =T)

@@

```{r}
hist(injuries$weight, breaks=10,main="Injuries by Weight", xlim=c(150,400),ylim=c(0,400), xlab="Weight",labels =T)
```

```{r}
hist(players$weight, breaks=10,main="Total Players Weight", xlim=c(150,400),ylim=c(0,400),  xlab="Weight",labels =T) 
```

@@

```{r}
hist(injuries$age, breaks=10,main="Injuries by Age", xlim=c(20,45),ylim=c(0,425), xlab="Age",labels =T)
```

```{r}
hist(players$age, breaks=10,main="Total Players Age", xlim=c(20,45),ylim=c(0,425),  xlab="Age",labels =T) 
```

@@

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

>>>>

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

@@

Over the past three decades, the NFL has come under increased scrutiny over the dangers of American football (@frontline-nfl-concussion). What initially began as concerns over the persistent brutality central to the sport, shifted into the more insidious issue of concussions and their role in the post-mortem diagnosis of Chronic Traumatic Encephalitic (CTE) (@mayoclinic-cte). The NFL has been making changes in the game in an attempt to prevent player injuries from rule modifications to investment in advancements to the players pads and helmets. Hopefully, the findings from this project can help inform players, coaches, and who knows, the NFL about the metrics for a higher risk in injury.

@@

Three data sets have been identified for this project. The primary data set ("Injury Player Data" contains 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season. Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics. The next is a "cumulative play data" containing information about every play ran in the 2019-2020 NFL season (42,186 observations). Information such as the teams playing, play description, and time in the game when it was ran. The last data set is the "player demographic data" containing information about all NFL players in the 2016-2019 seasons (11,145 observations). This data set has player metrics such as height, weight, date of birth, etc.

@@

@@

The primary idea behind the analysis for this project is to look at various parameters when an injury occurred and compare them to the rest of the NFL. If the parameters are related to the play itself, the parameter in the injury data set will be compared to cumulative play data set. Likewise, if the parameter is related to the injured player themselves, this parameter will be compared to the cumulative player data set. If there is a statistical difference between parameters, we can likely draw some conclusion about a parameter causing injuries.

@@

## Data aquisition
The "Injury Player Data" set was found from a Github repository. The repository belongs to a group of students who wanted to do a similar analysis looking at NFL injuries and various factors that could affect them.

https://github.com/sammieerne/NFL-Injury-Analysis/tree/main/Data

The "cumulative play data" set comes from a website called NFLsavant.com. This website is dedicated to providing NFL statistics to the public.

https://nflsavant.com/about.php

Lastly, the "player demographic data" set was also taken from a Github repository. Similar to the other repository, this one belongs to a group who wanted to analyze injuries in the NFL.

https://github.com/ericcrouse/NFL-Injury-Exploration/tree/main/data/raw

@@

As for the "cumulative play data" set, empty columns will be removed and 0's in the down column will be converted to "NA" values which indicate kickoffs & PATs. Lastly, a lot of unnecessary plays are included in the "cumulative play data" set such as end of quarter/games, 2-minute warnings, and timeouts so these observations can be filtered out.

@@

```{r}
ggplot(injuries, aes(x = qtr)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Injuries by Quarter", x = "Quarter", y = "Frequency")
```
```{r}
ggplot(pbp, aes(x = Quarter)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Total Plays by Quarter", x = "Quarter", y = "Frequency")
```

@@

```{r}
ggplot(injuries, aes(x = down)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Injuries by Down", x = "Down", y = "Frequency")
```

```{r}
ggplot(pbp, aes(x = Down)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Total Plays by Down", x = "Down", y = "Frequency")
```

@@

```{r}
ggplot(injuries, aes(x = height)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Injuries by Height", x = "Height", y = "Frequency")
```
```{r}
ggplot(players, aes(x = height)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Total Players by Height", x = "Height", y = "Frequency")
```

@@

```{r}
ggplot(injuries, aes(x = weight)) +
  geom_histogram(binwidth = 10, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Injuries by Weight", x = "Weight", y = "Frequency")
```

```{r}
ggplot(players, aes(x = weight)) +
  geom_histogram(binwidth = 10, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Total Players by Weight", x = "Weight", y = "Frequency")
```

@@

```{r}
ggplot(injuries, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Injuries by Age", x = "Age", y = "Frequency")
```

```{r}
ggplot(players, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
  labs(title = "Total Players by Age", x = "Age", y = "Frequency")
```

@@
